---
title: "Cincinnati crime"
author: "Rafal Nowicki"
date: "7 05 2020"
output: 
  html_document:
    code_folding: hide
    theme: journal
    css: style.css
---

```{r,include = FALSE}
library(RSocrata)
library(jsonlite)
library(tidyverse)
library(lubridate)
library(magrittr)
library(ggtext)
library(viridis)
library(RColorBrewer)
library(rayshader)
library(rgl)
library(htmltools)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(class.source = "graycode")

Sys.setlocale(category = "LC_ALL", locale = "english")
```

In this notebook, we use crime data from the City of Cincinnati, May 2019 to April 2020. "Incidents are the records, of reported crimes, collated by an agency for management."

Source: https://dev.socrata.com/foundry/data.cincinnati-oh.gov/k59e-2pvf

```{r, cache = TRUE}
cincinnati <- read.socrata("https://data.cincinnati-oh.gov/resource/k59e-2pvf.json?$where=date_reported between '2019-05-01' and '2020-04-30'")

#load("cincinnati_full.RData")

cincinnati %<>%
  mutate(
    datetime = as_datetime(date_reported)+2*60*60, # as_datetime converts date.. but than substracts 2 hours ?!
    date = date(datetime),
    hourofday = hour(datetime)
  )
```

## ggplot

The first task is to create a basic barplot for `offense` to get an idea of the prevalence of different types of crimes. It might make sense to flip the graph to make the labels more readable.

```{r, fig.align = "center"}
top_crimes <- cincinnati %>%
  group_by(offense) %>%
  summarise(count = n()) %>%
  filter(count > 500)


top_crimes %>%
  ggplot() +
  geom_col(aes(x = reorder(str_wrap(offense, 10), count), y = count), fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Number of crimes reported",
    x = "",
    y = "",
    subtitle = "by type, Cincinnati May 2019 to Apr 2020"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    text = element_text(size = 8),
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    plot.subtitle = element_text(color = "black", size = 10),
    plot.background = element_rect(fill = "darkgrey")
  )
```

Calculate the total number of criminal offenses grouped by `dst` and `offense`. Try to create boxplots for each `offense` that visualize the distribution of the total number of crimes. Again, flip the graph to make the labels more readable.

```{r, fig.align = "center"}
cincinnati %>%
  group_by(dst, offense) %>%
  summarise(total = n()) %>%
  filter(offense %in% top_crimes$offense) %>%
  ggplot(aes(x = reorder(str_wrap(offense, 10), total), y = total)) +
  geom_boxplot(
    outlier.colour = "red",
    outlier.size = 4,
    outlier.shape = 4,
    fill = "steelblue"
  ) +
  geom_jitter(
    shape = 4,
    position = position_jitter(0.2),
    col = "black",
    alpha = 0.5
  ) +
  labs(
    x = "",
    y = "",
    title = "Number of cases in districts",
    subtitle = "by type, Cincinnati May 2019 to Apr 2020"
  ) +
  coord_flip() +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    text = element_text(size = 8),
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    plot.subtitle = element_text(color = "black", size = 10),
    plot.background = element_rect(fill = "darkgrey")
  )
```

Create a plot that visualizes the total number of criminal offenses over `hourofday`. Draw separate lines for each `dayofweek` (note https://ggplot2.tidyverse.org/reference/aes_group_order.html).

```{r, fig.align = "center"}
cincinnati %>%
  group_by(dayofweek, hourofday) %>%
  summarise(count = n()) %>%
  drop_na() %>%
  ggplot() +
  geom_line(aes(x = hourofday, y = count, col = dayofweek), size = 1.05) +
  scale_colour_brewer(palette = "Set1") +
  #scale_colour_manual(values = c("#F8B2F8", "#ACA5F1", "#5598D9", "#0085B0", "#006E7C", "#005447", "#00391A"))+
  geom_smooth(
    aes(x = hourofday, y = count),
    col = "#CBCBCB",
    size = 1.3,
    alpha = 0.4,
    fill = "white",
    method = "gam"
  ) +
  labs(x = "", y = "", title = "**Offence cases on <span style='color:#377EB8;'>Mon</span>, <span style='color:#FFFF33;'>Tue</span>, <span style='color:#A65628;'>Wed</span>, <span style='color:#FF7F00;'>Thu</span>, <span style='color:#E41A1C;'>Fri</span>, <span style='color:#4DAF4A;'>Sat</span>, <span style='color:#984EA3'>Sun</span>** <br><span style = 'font-size:11pt'>by hour, Cincinnati May 2019 to Apr 2020</span> </br>") +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_markdown(lineheight = 1.1),
    legend.position = "none",
    plot.background = element_rect(fill = "darkgrey")
  )
```

Focus only on the two most common types of crimes. Visualize the total number of offenses for these types of crimes over `hourofday`, separate for each `dayofweek`. Stated differently -- find a good way to visualize crime trends with an additional grouping variable.

```{r, fig.align = "center"}
cincinnati %>%
  filter(offense %in% c("THEFT", "CRIMINAL DAMAGING/ENDANGERING")) %>%
  group_by(offense, dayofweek, hourofday) %>%
  summarise(count = n()) %>%
  drop_na() %>%
  ggplot() +
  geom_line(aes(x = hourofday, y = count, col = offense)) +
  facet_wrap( ~ dayofweek) +
  labs(
    title = "**Cases of
      <span style='color:#4682B4;'>Criminal Damagings</span> and
      <span style='color:#8B0000;'>Thefts</span>**
      <br><span style = 'font-size:11pt'>by day and hour, Cincinnati May 2019 to Apr 2020</span> </br>",
    x = "",
    y = ""
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot" ,
    legend.position  = "none",
    plot.title = element_markdown(lineheight = 1.1),
    plot.background = element_rect(fill = "darkgrey")
  ) +
  scale_colour_manual(values = c("steelblue", "darkred"))
```

Again, plot the total number of crimes over `hourofday`, but this time we want to have a separate plot for each `offense` (`THEFT` and `CRIMINAL DAMAGING/ENDANGERING`) and `dayofweek` combination. Add an additional layer of your choice to each individual plot that uses another `geom` (with the same data).

```{r, fig.align = "center"}
supp.labs <- c("CRIMINAL DAMAGING", "THEFT")
names(supp.labs) <- c("CRIMINAL DAMAGING/ENDANGERING", "THEFT")

cincinnati %>%
  filter(offense %in% c("THEFT", "CRIMINAL DAMAGING/ENDANGERING")) %>%
  group_by(offense, dayofweek, hourofday) %>%
  summarise(count = n()) %>%
  drop_na() %>%
  ggplot(aes(x = hourofday, y = count)) +
  geom_area(aes(fill = offense)) +
  facet_grid(offense ~ dayofweek, labeller = labeller(offense = supp.labs)) +
  labs(
    x = "",
    y = "",
    title = "Number of offenses",
    subtitle = "by day, hour and type, Cincinnati May 2019 to Apr 2020"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    plot.subtitle = element_text(color = "black", size = 10),
    plot.background = element_rect(fill = "darkgrey"),
    legend.position = "none"
  ) +
  scale_fill_manual(values = c("darkred", "steelblue"))
```

Try to create a heatmap (not a geographic map) that visualizes the total number of crimes per `hourofday` and `dayofweek`. Check the `ggplot()` documentation to find out which `geom_function` can be used for this task.

```{r, fig.align = "center"}
cincinnati %>%
  mutate(dayofweek = as_factor(dayofweek)) %>%
  mutate(
    dayofweek = fct_relevel(
      dayofweek,
      "MONDAY",
      "TUESDAY",
      "WEDNESDAY",
      "THURSDAY",
      "FRIDAY",
      "SATURDAY",
      "SUNDAY"
    )
  ) %>%
  group_by(dayofweek, hourofday) %>%
  summarise(count = n()) %>%
  drop_na() %>%
  ggplot() +
  geom_tile(aes(x = hourofday, y = dayofweek, fill = count), col = "black") +
  scale_fill_viridis_c("Total\nCases", option = "C") +
  labs(
    x = "",
    y = "",
    title = "Number of crimes reported",
    subtitle = "by day and hour, Cincinnati May 2019 to Apr 2020"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.background = element_rect(fill = "darkgrey"),
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    plot.subtitle = element_text(color = "black", size = 10)
  )
```

Extend your heatmap by adding more information. This could e.g. be a `geom_text` layer that prints the total number of crimes in each cell of the heatmap. 

```{r, fig.align = "center"}
# pl <- cincinnati %>%
#   mutate(dayofweek = as_factor(dayofweek)) %>%
#   mutate(
#     dayofweek = fct_relevel(
#       dayofweek,
#       "MONDAY",
#       "TUESDAY",
#       "WEDNESDAY",
#       "THURSDAY",
#       "FRIDAY",
#       "SATURDAY",
#       "SUNDAY"
#     )
#   ) %>%
#   group_by(dayofweek, hourofday) %>%
#   summarise(count = n()) %>%
#   drop_na() %>%
#   ggplot() +
#   geom_tile(aes(x = hourofday, y = dayofweek, fill = count), col = "black") +
#   scale_fill_viridis_c("Total\nCases", option = "C") +
#   labs(
#     x = "",
#     y = "",
#     title = "Number of crimes commited",
#     subtitle = "by day and hour, Cincinnati May 2019 to Apr 2020"
#   )
# 
# 
#  plot_gg(
#    pl,
#   width = 5,
#  height = 4,
#  scale = 300,
#  multicore = TRUE,
#  fov = 70,
#  zoom = 0.45,
#  theta = 330,
#  phi = 40,
#  background = "#A9A9A9",
#  shadowcolor = "#3a4f70"
#  )
#
##############################################################
#  filename <- writeWebGL(dir = file.path(tempdir(), "webGL"),
#                        width = 600, reuse = TRUE)
################################################################

# #render_snapshot(clear = TRUE)
#rglwidget(width = 400, height = 800)
#render_depth(focus = 0.68, focallength = 200,clear=TRUE)


htmltools::includeHTML("plot3d.html")
```

<br>
Building on the previous examples, create two separate heatmaps for `THEFT` and `CRIMINAL DAMAGING/ENDANGERING` via faceting.

```{r, fig.align = "center"}
cincinnati %>%
  group_by(offense, dayofweek, hourofday) %>%
  filter(offense %in% c("THEFT", "CRIMINAL DAMAGING/ENDANGERING")) %>%
  summarise(count = n()) %>%
  drop_na() %>%
  ggplot() +
  geom_tile(aes(x = hourofday, y = dayofweek, fill = count), col = "black") +
  facet_wrap(. ~ offense) +
  scale_fill_viridis_c("Total\nCases", option = "E") +
  labs(
    x = "",
    y = "",
    title = "Number of crimes reported",
    subtitle = "by day and hour, Cincinnati May 2019 to Apr 2020"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.background = element_rect(fill = "darkgrey"),
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    plot.subtitle = element_text(color = "black", size = 10)
  )
```


Finally, build a plot that visualizes the crime trend/ the total number of crime offenses over the whole time range of our data. Add any layer, aesthetic or facet that you find useful. 

```{r, fig.align = "center"}
cincinnati %>%
  transmute(
    date = as_datetime(date_reported) + 2 * 60 * 60,
    day = day(date),
    month = fct_recode(
      as.factor(month(date)),
      "May 19" = "5",
      "Jun 19" = "6",
      "Jul 19" = "7",
      "Aug 19" = "8",
      "Sep 19" = "9",
      "Oct 19" = "10",
      "Nov 19" = "11",
      "Dec 19" = "12",
      "Jan 20" = "1",
      "Feb 20" = "2",
      "Mar 20" = "3",
      "Apr 20" = "4"
    )
  ) %>%
  mutate(
    month = fct_relevel(
      month,
      "May 19",
      "Jun 19",
      "Jul 19",
      "Aug 19",
      "Sep 19",
      "Oct 19",
      "Nov 19",
      "Dec 19",
      "Jan 20",
      "Feb 20",
      "Mar 20",
      "Apr 20"
    )
  ) %>%
  group_by(month, day) %>%
  summarise(count = n()) %>%
  drop_na() %>%
  ggplot() +
  geom_tile(aes(x = day, y = month, fill = count), col = "black") +
  scale_fill_viridis_c("Total\nCases", option = "E") +
  labs(
    x = "",
    y = "",
    title = "Reported offense cases daily",
    subtitle = "Cincinnati, May 2019 to Apr 2020"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.background = element_rect(fill = "darkgrey"),
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    plot.subtitle = element_text(color = "black", size = 10)
  )

```

Classic way of showing time series data:

```{r, fig.align= "center"}
cincinnati %>%
  group_by(date) %>%
  summarise(n = n()) %>%
  drop_na() %>%
  ggplot(aes(x = date, n)) +
  geom_line(col = "#1B305B", size = 1.03) +
  geom_smooth(method = "gam", col = "#FFEA46", fill = "darkred") +
  labs(
    x = "",
    y = "",
    title = "Reported offense cases",
    subtitle = "Cincinnati, May 19 to Apr 20"
  ) +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.background = element_rect(fill = "darkgrey"),
    plot.title = element_text(color = "black", size = 12, face = "bold"),
    plot.subtitle = element_text(color = "black", size = 10)
  )

```